Exploring Constraint Grammar as a SAT problem

### 2015-03-27 paper/nodalida2015.tex
stuff

\begin{verbatim}
"<la>"
        "el" det def f sg
        "lo" prn pro p3 f sg
"<casa>"
        "casa" n f sg
        "casar" vblex pri p3 sg
\end{verbatim}

@@

\begin{itemize}
\item [] \texttt{REMOVE prn IF (1 n) ;} \\
             \texttt{REMOVE det IF (1 vblex) ;}
\end{itemize}

Standard CG will apply one\footnote{To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, 
neither of the rules would be applied.} of the rules to the word \emph{la}; 
either the one that comes first, or by some other heuristic. 
The second rule will not fire, because it would remove the last reading. 
Both readings of \emph{casa} are left untouched by these rules.

@@

%If we add a word, \emph{la casa grande} `the big house' 
%and a rule that removes verb if followed by an adjective
%\texttt{REMOVE (vblex) IF (1 (adj)) ;}
%we will get a unique solution with \texttt{det n adj}.
%Standard CG will also remove the verb reading, 
%but the choice of the word \emph{la} doesn't contribute to

@@

\begin{itemize}
\item [] \texttt{REMOVE n IF (-1 prn) ;} \\
             \texttt{REMOVE vblex IF (-1 det) ;}
\end{itemize}

@@

To replace the sequential order, we have experimented with different strategies:
\begin{itemize}
\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Didn't give the best results.
\item Maximise number of instances when a rule is applied---\todo{comment about this too}
\end{itemize}

@@

\paragraph{Performance against VISL CG3}
We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg. Table~\ref{table:results} shows how much the results agree with the result given by VISL CG3 parser with the same grammar and text.

>>>>

\begin{verbatim}
"<la>"
        "el" det def f sg
        "lo" prn pro p3 f sg
"<casa>"
        "casa" n f sg
        "casar" v pri p3 sg
        "casar" v imp p2 sg
\end{verbatim}

@@

\begin{itemize}
\item [] \texttt{REMOVE prn IF (1 n) ;} \\
             \texttt{REMOVE det IF (1 v) ;}
\end{itemize}

Standard CG will apply one\footnote{To be more cautious, we could require the word at position +1 be disambiguated fully (\texttt{1C} instead of \texttt{1}), but in that case, 
neither of the rules would be applied.} of the rules to the word \emph{la}; 
either the one that comes first, or by some other heuristic. 
The second rule will not fire, because it would remove the last reading. 
All readings of \emph{casa} are left untouched by these rules.

@@

%If we add a word, \emph{la casa grande} `the big house' 
%and a rule that removes verb if followed by an adjective
%\texttt{REMOVE (v) IF (1 (adj)) ;}
%we will get a unique solution with \texttt{det n adj}.
%Standard CG will also remove the verb reading, 
%but the choice of the word \emph{la} doesn't contribute to

@@

\begin{itemize}
\item [] \texttt{REMOVE n IF (-1 prn) ;} \\
             \texttt{REMOVE v IF (-1 det) ;}
\end{itemize}

@@

To replace the sequential order, we have experimented with different strategies:
\begin{itemize}
\item Maximise number of rules used---assume if someone has written a rule, it is worth applying. Downside is that if a rule leads to a conflict in one case, it won't be applied anywhere.
\item Maximise number of instances when a rule is applied---This gave better results. Assume that whenever we can apply a rule, we should do it.
\item Maximise number of true analyses---With this heuristic, the solution \texttt{prn v} would be chosen, because there happens to be two analyses with \texttt{v}, and choosing that, 3 analyses are true, since there is no rule to disambiguate between imperative and indicative. We want analyses to be true by default
\end{itemize}

@@

\paragraph{Performance against VISL CG3}

\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg. Table~\ref{table:results} shows how much the results agree with the result given by VISL CG3 parser with the same grammar and text.
